### 2017-08-21 README.Rmd
major structuring

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Maragra and malaria

_Joe Brew, Elisa Sicuri, Laia Cirera, Kizito Gondo, Elton Dorkin_

# Details

- This paper-as-a-package concept was inspired by https://github.com/Pakillo/template