### 2017-08-28 inst/rmd/maragra_preliminary_analysis.Rmd
done with preliminary analysis skeleton

---
date: ""
fig_height: 2.6
fig_width: 4
output:
  html_document:
    css: resources/style.css
    pandoc_args: [
      "+RTS", "-K16000m",
      "-RTS"
    ]
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# Packages
library(tidyverse)
library(knitr)
library(brew)
library(maragra)
library(knitr)
library(googleVis)
op <- options(gvis.plot.tag='chart')

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

<p style="text-align:left;">
[www.economicsofmalaria.com](www.economicsofmalaria.com)
<span style="float:right;">joebrew@gmail.com</span>
</p>

```{r}
# Read in data
ab <- maragra::ab
ab_panel <- maragra::ab_panel
clinic <- maragra::clinic
clinic_agg <- maragra::clinic_agg
mc <- maragra::mc
workers <- maragra::workers
```

# Maragra and Malaria: preliminary analysis {.tabset}

This document contains some basic preliminary analysis of data relating to malaria, IRS, and absenteeism at Maragra Açucar, in Maragra, Mozambique. To explore a specific section, click any of the below tabs.

## Description of data

```{r}
d <- data(package = "maragra")
## assign it to use later
nm <- d$results[, "Item"]
```

The "Maragra database" consists of the following `r length(nm)` tables:

```{r}
## call the promised data
data(list = nm, package = "maragra")
## get the dimensions of each data set
x <- lapply(mget(nm), dim)
results <- list()
for (i in 1:length(x)){
  z <- x[[i]]
  results[[i]] <- data_frame(`Table name` = names(x)[i],
             `Number or rows` = z[1],
             `Number of columns` = z[2])
}
results <- bind_rows(results)

kable(results)
```

These datasets are all readily available in the (private) `maragra` R package (access credentials given to collaborators upon [joebrew@gmail.com](request)). All datasets except for `ab_panel` are _as is_ (ie, in their original raw format), except for 4 kinds of changes:

1. Removal of unecessary information (columns deemed irrelevant to the study have been removed).  
2. Modification of column names (all data now employ underscores instead of spaces, and are all lowercase).    
3. Standardization of data formats (all dates are now in a YYYY-MM-DD format, all names use similar capitalization, irrelevant tablet-created timestamps, etc. have been removed).  
4. "Feature generation": certain columns were created in order to ease grouped analysis and merging of datasets (for example, `year_month`, `day_number`, etc.).

The `ab_panel` dataset is _not_ raw data; rather, it is an amolgamation of the `ab` and `workers` datasets, using absence data from the former with worker elibility dates from the latter, so as to create a "panel" style dataset (ie, one row for each day in which a worker was _supposed_ to work).

The `clinic` and `clinic_agg` are similar, but not identical, in form. The former has individual-level data (useful for pairing with absences and demographic census data), whereas the latter is simply the raw counts of cases by nationality over time. Though `clinic` is much richer and more detailed than `clinic_agg`, is only covers the period from 2014 through 2016, whereas the latter covers a larger time period (2010-2016).

For a "deep dive" into each kind of data, click the tabs at the top of this page.

In order to reproduce this analysis (and others related to these datasets), follow the instructions on this project's [research compendium page](https://github.com/joebrew/maragra#purpose).

## IRS analysis

Here is a chart.

```{r}
ggplot(data = data.frame(a = 1:10,
                         b = 1:10)) +
  geom_point(aes(x = a, y = b))
```

## Absenteeism analysis

Here is a chart.

```{r}
ggplot(data = data.frame(a = 1:10,
                         b = 1:10)) +
  geom_point(aes(x = a, y = b))
```

## Clinic analysis

Here is a chart.

```{r}
ggplot(data = data.frame(a = 1:10,
                         b = 1:10)) +
  geom_point(aes(x = a, y = b))
```

## Next steps