### 2017-10-24 inst/rmd/ihmt_presentation_brew_2017.Rmd
major presentation update

---
title: !r "Effectiveness of private sector malaria control: The case of sugarcane workers in Mozambique"
subtitle: IHMT, October, 2017, Lisboa, Portugal
fig_height: 5
fig_width: 7
bibliography: bibliography.bib
output:
  beamer_presentation:
    includes:
      in_header: preamble_beamer.sty
    theme: "Berlin"
    colortheme: "seagull"
    fonttheme: "structuresmallcapsserif"
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F, results = 'hide'}
# All beamer themes at http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme_and_color.html
library(knitr)
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)

# Packages
library(ggplot2)
library(cism)
library(rgeos)
library(maptools)
library(rgdal)
library(tidyr)
library(RColorBrewer)
library(dplyr)
library(leaflet)
library(readr)
library(ggthemes)
library(ggrepel)
library(knitr)
options(knitr.table.format = "latex")
library(maragra)
```

```{r, results = 'asis'}
# Print the author names and affiliations
x <- make_authors(include_address = FALSE, include_country = TRUE, seperator = ' \n\n ', cat_it = FALSE)
cat(x)
```

# Introduction

## Context

- Malaria has a nearly unquantifiably large economic impact.
- Many channels: fertility, fecundity, saving, investment [@Shretta2016], risk perception, productivity, absenteeism, human capital accumulation [@CastelBranco2014], mortality, costs of care [@Sachs2002].
- Cost-benefit studies often only consider the costs of an intervention and associated costs of care, without quantifying the societal cost of non-intervention.

## What we already know

- Malaria is associated with absenteeism in workers [@Nonvignon_2016].
- Malaria has a negative effect on GDP [@Orem_2012] and growth [@McCarthy_2000].
- Malaria control is cost-effective from the societal/public perspective [@Purdy_2013].
- Indoor residual spraying (IRS) is cost-effective [@Howard_2017], [@White_2011].

## What we want to know

What is the investment case from the private perspective?

- From a purely financial perspective, what benefits does a private company experience?  
- What is the short-term cost-effectiveness of IRS for large companies in malaria-endemic regions?
- Is malaria control just good "corporate social responsibility"? Or is it also good business?

## Research questions

We can't answer all the previous questions (yet). So we focus on one:

**What is the short-term effect on absenteeism of IRS at the homes of sugarcane workers?**

# Methods

## Reproducibility

- **Code:** [github.com/joebrew/maragra](https://github.com/joebrew/maragra)
- **Data:** Only available to authorized collaborators, but "dummy data" has been created to validate methods  
- **Code bundling:** All code is in an R package: `devtools::install_github('joebrew/maragra)`

## Identification strategy

Some text here.

# Results

## Reduction in absenteeism

Some text here.

## Reduction in illness

Some text here.

# Discussion

## General

## Implications

## Limitations

## Thank you

**Email:** joe@economicsofmalaria.com

**Presentation:** [economicsofmalaria.com/ihmt](http://economicsofmalaria.com/ihmt)

## References