### 2017-10-29 inst/rmd/ihmt_presentation_brew_2017.Rmd
done with ihmt presentation

---
title: "Effectiveness of private sector malaria control: The case of sugarcane workers in Mozambique"
subtitle: 'IHMT, October, 2017, Lisboa, Portugal'
fig_height: 5
fig_width: 7
bibliography: bibliography.bib
output:
  beamer_presentation:
    includes:
      in_header: preamble_beamer.sty
    theme: "Berlin"
    colortheme: "seagull"
    fonttheme: "structuresmallcapsserif"
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F, results = 'hide'}
# All beamer themes at http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme_and_color.html
library(knitr)
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
opts_chunk$set(dev = 'pdf')

# Packages
library(ggplot2)
library(cism)
library(rgeos)
library(maptools)
library(rgdal)
library(tidyr)
library(RColorBrewer)
library(dplyr)
library(leaflet)
library(readr)
library(ggthemes)
library(ggrepel)
library(knitr)
options(knitr.table.format = "latex")
library(maragra)
```

```{r, results = 'asis'}
# Print the author names and affiliations
x <- make_authors(include_address = FALSE, include_country = TRUE, seperator = ' \n\n ', cat_it = FALSE)
cat(x)
```

```{r}
# Read in data
ab <- maragra::ab
ab_panel <- maragra::ab_panel
bairros <- maragra::bairros
census <- maragra::census
clinic <- maragra::clinic
clinic_agg <- maragra::clinic_agg
irs <- maragra::irs
mc <- maragra::mc
workers <- maragra::workers

# Define function fro breaking the days_since var
break_days_since <- function(days_since){
  out <- ifelse(days_since > 270, '270+',
    ifelse(days_since > 180, '181-270',
           ifelse(days_since > 90, '091-180',
                  ifelse(days_since > 60, '061-090',
                         ifelse(days_since > 30, '031-060',
                                ifelse(days_since >= 0, '000-030',
                                       ifelse(days_since < 0, 'Before', NA)))))))
  out <- factor(out,
                levels = c('Before',
                           '000-030',
                           '031-060',
                           '061-090',
                           '091-180',
                           '181-270',
                           '270+'))
  return(out)
}

# Define function for making season
make_season <- function(date){
  out <- ifelse(as.numeric(format(date, '%m')) %in% c(11:12, 1:3),
                   'Malaria season', 'Off season')
  out <- factor(out, levels = c('Off season', 'Malaria season'))
  return(out)
}

library(MatchIt)
set.seed(1234)

# Create matching data for propensity scoring
right_side <- irs %>%
  group_by(unidade) %>% 
  tally %>%
  ungroup %>%
  mutate(received = ifelse(n > 0, TRUE, FALSE))

left_side <-
  workers %>%
  dplyr::select(oracle_number, unidade, permanent_or_temporary, department, sex, 
                date_of_birth,
                perm_id) %>%
  # bring in census address
  left_join(census %>%
              dplyr::select(perm_id,
                            longitude,
                            latitude,
                            maragra_bairro,
                            maragra_fabrica))

psm <- left_join(x = left_side,
                 y = right_side,
                 by = 'unidade') %>%
  # Keep only those who are censed
  filter(!is.na(longitude),
         !is.na(latitude)) %>%
  # Keep only those who live within maragra
  filter(maragra_fabrica |maragra_bairro) %>%
  mutate(received = ifelse(is.na(received), FALSE, received)) %>%
  mutate(age = round((as.numeric(as.Date('2016-01-01') - date_of_birth)) / 365.25, digits = 2)) %>%
  filter(!is.na(sex),
         !is.na(age)) %>%
  dplyr::select(-unidade, -date_of_birth, -n)

psmt1 <- psm
names(psmt1) <- toupper(names(psmt1))
psmt1 <- psmt1 %>%
  mutate(RECEIVED = ifelse(RECEIVED == TRUE,'IRS', 'No IRS')) %>%
  mutate(STATUS = PERMANENT_OR_TEMPORARY)

pacman::p_load(tableone)
table1 <- CreateTableOne(vars = toupper(c('STATUS', 'department', 'age',
                                  'sex', 'received')), 
                         data = psmt1, 
                         factorVars = toupper(c('STATUS', 'department',
                                  'sex')), 
                         strata = 'RECEIVED')
table1 <- print(table1, 
                printToggle = FALSE, 
                noSpaces = TRUE)

match.it <- matchit(received ~ age + sex + permanent_or_temporary + department, data = psm, method="nearest", ratio=1)
# Save matched samples
matched <- match.data(match.it)[1:ncol(psm)]
a <- summary(match.it)

# Create model data
model_data_propensity <-
  ab_panel %>%
  filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs) %>%
  left_join(matched) %>%
  mutate(days_since = break_days_since(days_since)) %>%
  filter(!is.na(days_since)) %>%
  # mutate(time_period = lendable::time_period_extract(date)) %>%
  # mutate(time_period = as.character(time_period)) %>%
  mutate(time_period = make_season(date = date)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))

# Create model data for all obs (not just propensity score)
model_data <-
  ab_panel %>%
  # filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs) %>%
  # left_join(matched) %>%
  mutate(days_since = break_days_since(days_since)) %>%
  filter(!is.na(days_since)) %>%
  # mutate(time_period = lendable::time_period_extract(date)) %>%
  # mutate(time_period = as.character(time_period)) %>%
  mutate(time_period = make_season(date = date)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick)) #%>%
  # # remove the period 30 days after spraying
  # filter(days_since >= 30 | days_since <= 0)
# Need to add seasonality
# Need to adjust for houses on and off site

fit_propensity <- glm(absent ~ days_since * time_period,
           family = binomial('logit'),
          data = model_data_propensity)
x <- exp(coef(fit_propensity))
ors_propensity <- exp(confint(fit_propensity))
ors_propensity <- data.frame(ors_propensity)
names(ors_propensity) <- c('Lower', 'Upper')

ors_propensity <- cbind(x, ors_propensity)
ors_propensity$Variable <- row.names(ors_propensity)
ors_propensity <- ors_propensity %>%
  dplyr::rename(OR = x) %>%
  dplyr::select(Variable, OR, Lower, Upper) %>%
  mutate(OR = round(OR, digits = 3),
         Lower = round(Lower, digits = 3),
         Upper = round(Upper, digits = 3))

# Need to add seasonality
# Need to adjust for houses on and off site
fit_propensity_sick <- glm(absent_sick ~ days_since * time_period,
           family = binomial('logit'),
          data = model_data_propensity)
x <- exp(coef(fit_propensity_sick))
ors_propensity_sick <- exp(confint(fit_propensity_sick))
ors_propensity_sick <- data.frame(ors_propensity_sick)
names(ors_propensity_sick) <- c('Lower', 'Upper')

ors_propensity_sick <- cbind(x, ors_propensity_sick)
ors_propensity_sick$Variable <- row.names(ors_propensity_sick)
ors_propensity_sick <- ors_propensity_sick %>%
  dplyr::rename(OR = x) %>%
  dplyr::select(Variable, OR, Lower, Upper) %>%
  mutate(OR = round(OR, digits = 3),
         Lower = round(Lower, digits = 3),
         Upper = round(Upper, digits = 3))

# Need to adjust for houses on and off site
fit <- glm(absent ~ days_since * time_period,
           family = binomial('logit'),
          data = model_data)
x <- exp(coef(fit))
ors <- exp(confint(fit))
ors <- data.frame(ors)
names(ors) <- c('Lower', 'Upper')

ors <- cbind(x, ors)
ors$Variable <- row.names(ors)
ors <- ors %>%
  dplyr::rename(OR = x) %>%
  dplyr::select(Variable, OR, Lower, Upper) %>%
  mutate(OR = round(OR, digits = 3),
         Lower = round(Lower, digits = 3),
         Upper = round(Upper, digits = 3))

# Need to add seasonality
# Need to adjust for houses on and off site
fit_sick <- glm(absent_sick ~ days_since * time_period,
           family = binomial('logit'),
          data = model_data)
x <- exp(coef(fit_sick))
ors_sick <- exp(confint(fit_sick))
ors_sick <- data.frame(ors_sick)
names(ors_sick) <- c('Lower', 'Upper')

ors_sick <- cbind(x, ors_sick)
ors_sick$Variable <- row.names(ors_sick)
ors_sick <- ors_sick %>%
  dplyr::rename(OR = x) %>%
  dplyr::select(Variable, OR, Lower, Upper) %>%
  mutate(OR = round(OR, digits = 3),
         Lower = round(Lower, digits = 3),
         Upper = round(Upper, digits = 3))

```

# Introduction

## Context

- Malaria has a nearly unquantifiably large economic impact.
- Many channels: fertility, fecundity, saving, investment [@Shretta2016], risk perception, productivity, absenteeism, human capital accumulation [@CastelBranco2014], mortality, costs of care [@Sachs2002].
- Cost-benefit studies often only consider the costs of an intervention and associated costs of care, without quantifying the societal cost of non-intervention.

## What we already know

- Malaria is associated with absenteeism in workers [@Nonvignon_2016].
- Malaria has a negative effect on GDP [@Orem_2012] and growth [@McCarthy_2000].
- Malaria control is cost-effective from the societal/public perspective [@Purdy_2013].
- Indoor residual spraying (IRS) is cost-effective [@Howard_2017], [@White_2011].

## What we want to know

What is the investment case _from the investor's perspective_?

- Is malaria control just good "corporate social responsibility"? Or is it also good business?
- From a purely financial/investment point-of-view, what benefits does a private company experience in engaging in malaria control?  
- What is the short-term **benefits** of IRS for large companies in malaria-endemic regions?
- What are the **costs** of of carrying out IRS for large companies?

## Research questions

We can't answer all the previous questions (yet). So we focus on one:

**What is the short-term effect of IRS on worker absenteeism and clinical illness among sugarcane workers?**

## Research site

```{r}
par(mfrow = c(2,2))
plot(cism::africa,
     col = 'grey',
     border = 'white',
     lwd = 0.1)
title('Africa')
for(i in (seq(0.5, 1.5, length = 2))^1.3){
  points(32.779401, -25.452049, col = adjustcolor('red', alpha.f = 0.5), cex = i)
}
plot(cism::moz2,
     col = 'grey',
     border = 'white',
     lwd = 0.1)
title('Mozambique')
for(i in (seq(0.5, 2.5, length = 3))^1.3){
  points(32.779401, -25.452049, col = adjustcolor('red', alpha.f = 0.5), cex = i)
}
man <- cism::man3
plot(man,
     col = 'grey', 
     border = 'white',
     lwd = 0.1)
title('Manhiça district')
for(i in (seq(0.5, 3, length = 4))^1.3){
  points(32.779401, -25.452049, col = adjustcolor('red', alpha.f = 0.5), cex = i)
}
plot(man[man@data$NAME_3 %in% c('Manhica - Sede', 'Maluana'),],
     col = 'grey', 
     border = 'white',
     lwd = 0.1)
title('Manhiça and Maluana posts')
for(i in (seq(0.5, 3.5, length = 4))^1.3){
  points(32.779401, -25.452049, col = adjustcolor('red', alpha.f = 0.5), cex = i)
}
par(mfrow = c(1,1))
```

## Research Site II

\includegraphics{images/sat}

# Methods

## Identification strategy

- 1 intervention (IRS, time to/from)
- 2 outcomes (absence and illness, probabilistic / binomial)
- Many confounders (age, worker type, seasonality, etc.).

\begin{small}
\begin{equation}
\operatorname{Pr}(\text{Outcome} = 1 \mid \text{X}) = \beta_{0} + \beta_{1} \text{Location} + \beta_{2} \text{Season} + (\beta_3{IRS}*\beta_4{IRS_t} + ... )
\end{equation}
\end{small}

## Modeling

We employ two approaches:

1. **Propensity score matching** of workers who _ever_ received IRS with workers who _never_ received IRS. **Advantage**: No need to adjust for confounders with a matched sample, thereby avoiding reduction in degrees of freedom

2. **Regresion-discontinuity** of only those workers who _ever_ received IRS (ie, ignoring those who never received IRS). **Advantage**: Those who _never_ received IRS may be qualitatively different, and therefore not an appropriate comparison group.

## Propensity score matching {.allowframebreaks}

- We generate a matched sample of similar workers by first estimating the likelihood of having ever received the intervention, given a worker's age, sex, department and temporary vs. permanent status. 
 - This is necessary due to below, significant differences:

\begin{tiny}

```{r}
kable(table1[,1:3],  
      align = 'c', 
      caption = 'Comparison of unmatched samples') %>%
  kableExtra::kable_styling(full_width = FALSE)
```

\end{tiny}

We match, employing the nearest neighbor method for identifying those workers from our control group who most resemble those workers in the treatment group. [@Ho_Imai_King_Stuart_2007].

- Our match is a 1-to-1 cut, meaning those control workers who do not resemble those in the treatment group are left out of primary analysis. The below table shows the match results.

\begin{tiny}

```{r, results = 'asis'}
kable(a$nn, digits = 2, align = 'c', 
      caption = 'Sample sizes')
```

\end{tiny}

The distributions of our numeric variables are now extremely similar:

\begin{tiny}

```{r, results = 'asis'}
kable(a$sum.matched[c(1,2,3,4)], digits = 2, align = 'c', 
      caption = 'Summary of balance for matched data') 
```

\end{tiny}

## Regression discontunity analysis

- We simply only consider those workers who _ever_ got IRS.
- We take into account one full year prior to IRS and one full year after IRS.
- Our dataset constitutes one observation per worker-day. 
- We incidentally achieve a sort of "matching" through the fact that workers are their own controls.

# Results

## Descriptive: absenteeism by time from/to intervention

```{r}
plot_data <-
  ab_panel %>%
  # filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs) %>%
  mutate(months_since = days_since %/% 30) %>%
  filter(!is.na(months_since)) %>%
  # mutate(time_period = lendable::time_period_extract(date)) %>%
  # mutate(time_period = as.character(time_period)) %>%
  mutate(time_period = make_season(date = date)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))

x <- plot_data %>%
  group_by(months_since) %>%
  summarise(absences = length(which(absent)),
            sick_absences = length(which(absent_sick)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100)

ggplot(data = x,
       aes(x = months_since,
           y = absenteeism_rate)) +
  geom_step(color = 'darkred', alpha = 0.8) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate',
       title = 'Before/after IRS',
       subtitle = 'All absences') +
  theme_maragra()
```

## Descriptive: absenteeism by time from/to intervention (with local regression lines)

```{r}
plot_data <-
  ab_panel %>%
  # filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs) %>%
  mutate(months_since = days_since %/% 30) %>%
  filter(!is.na(months_since)) %>%
  # mutate(time_period = lendable::time_period_extract(date)) %>%
  # mutate(time_period = as.character(time_period)) %>%
  mutate(time_period = make_season(date = date)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))

x <- plot_data %>%
  group_by(months_since) %>%
  summarise(absences = length(which(absent)),
            sick_absences = length(which(absent_sick)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100) %>%
  mutate(before = months_since < 0)

ggplot(data = x,
       aes(x = months_since,
           y = absenteeism_rate,
           group = before)) +
  geom_smooth() +
  geom_step(color = 'darkred', alpha = 0.8) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate',
       title = 'Before/after IRS',
       subtitle = 'All absences') +
  theme_maragra()
```

## Descriptive: sick absenteeism by time from/to intervention

```{r}
ggplot(data = x,
       aes(x = months_since,
           y = sick_absenteeism_rate,
           group = before)) +
  geom_smooth() +
  geom_step(color = 'darkred', alpha = 0.8) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate',
       title = 'Before/after IRS',
       subtitle = 'Sick absences only') +
  theme_maragra()
```

## Descriptive: absenteeism by time from/to intervention (by time_period)

```{r}
plot_data <-
  ab_panel %>%
  # filter(oracle_number %in% matched$oracle_number) %>%
  left_join(irs) %>%
  mutate(months_since = days_since %/% 30) %>%
  filter(!is.na(months_since)) %>%
  # mutate(time_period = lendable::time_period_extract(date)) %>%
  # mutate(time_period = as.character(time_period)) %>%
  mutate(time_period = make_season(date = date)) %>%
  mutate(absent_sick = ifelse(is.na(absent_sick), FALSE, absent_sick))

x <- plot_data %>%
  group_by(months_since, time_period) %>%
  summarise(absences = length(which(absent)),
            sick_absences = length(which(absent_sick)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) %>%
  mutate(sick_absenteeism_rate = sick_absences / eligibles * 100) %>%
  mutate(before = months_since < 0)

ggplot(data = x %>% mutate(time_period = paste0(time_period)),
       aes(x = months_since,
           y = absenteeism_rate)) +
  geom_step(color = 'darkred', alpha = 0.8) +
  facet_wrap(~time_period, ncol = 1) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate') +
  theme_maragra()
```

## Same chart with local regression lines

```{r}
ggplot(data = x %>% mutate(time_period = paste0(time_period)),
       aes(x = months_since,
           y = absenteeism_rate,
           group = before)) +
  geom_smooth() +
  geom_step(color = 'darkred', alpha = 0.8) +
  facet_wrap(~time_period, ncol = 1) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate') +
  theme_maragra()
```

## Descriptive: sick absenteeism by time from/to intervention (by time_period)

```{r}

ggplot(data = x %>% mutate(time_period = paste0(time_period)),
       aes(x = months_since,
           y = sick_absenteeism_rate,
           group = before)) +
  geom_smooth() +
  geom_step(color = 'darkred', alpha = 0.8) +
  facet_wrap(~time_period, ncol = 1) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.5) +
  labs(x = 'Months relative to IRS',
       y = 'Absenteeism rate',
       subtitle = 'Sick absences only') +
  theme_maragra()
```

## Descriptive: absenteeism in binned time period

```{r}
x <- model_data %>%
  group_by(days_since) %>%
  summarise(absences = length(which(absent)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100)

ggplot(data = x,
       aes(x = days_since,
           y = absenteeism_rate)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           fill = 'darkblue') +
  theme_maragra() +
  labs(x = 'Days since IRS',
       y = 'Absenteeism rate',
       title = 'Raw absenteeism as a function of days since IRS') +
  geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%')))
```

## Descriptive: absenteeism in binned time periods by season

```{r}
x <- model_data %>%
  group_by(days_since, time_period) %>%
  summarise(absences = length(which(absent)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100)

ggplot(data = x %>% mutate(time_period = paste0(time_period)),
       aes(x = days_since,
           y = absenteeism_rate)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           fill = 'darkblue') +
  theme_maragra() +
  labs(x = 'Days since IRS',
       y = 'Absenteeism rate',
       title = 'Raw absenteeism as a function of days since IRS by time_period') +
  # geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%'))) +
  facet_wrap(~time_period, ncol = 1) +
  theme(axis.text.x = element_text(angle = 90))
```

## Modeling after matching {.allowframebreaks}

- Matched sample of `r nrow(matched)` workers, of which 50% received IRS and 50% did not.   
- Model only takes into account seasonality, since matching theoretically handles other differences.  
- For the purposes of this first pass, we "bin" IRS exposure and estimate a logit model to calculate odds ratios.

\vfill
\newpage

All absence:

\begin{tiny}

```{r, results = 'asis'}
# Combine into a df
kable(ors_propensity) 
```

\end{tiny}

\vfill
\newpage

```{r}
visualize_ors <- function(ors_object){
  ors_object <- ors_object %>% 
    filter(grepl('days_since', Variable),
           grepl('time_periodMalaria', Variable))
  ors_object <- ors_object %>%
    mutate(Variable = gsub('days_since', '', Variable),
           Variable = gsub(':time_periodMalaria season', '', Variable))
  ggplot(data = ors_object,
         aes(x = Variable,Variable,
             y = OR)) +
    geom_point(alpha = 0.7) +
    geom_errorbar(aes(ymin = Lower,
                       ymax = Upper),
                  alpha = 0.6,
                  size = 0.5) +
    geom_hline(yintercept = 1, lty = 2, alpha = 0.6, color = 'red') +
    labs(x = 'Days since IRS',
         y = 'OR (relative to pre-IRS period)',
         title = 'ORs during malaria season') +
    theme_maragra() 
}
visualize_ors(ors_object = ors_propensity)
```

\vfill
\newpage

Sick absence only:

\begin{tiny}

```{r}
kable(ors_propensity_sick) 
```

\end{tiny}

## Ever IRS'ers compared with never-IRS'ers

```{r}
x <- model_data %>%
  group_by(days_since) %>%
  summarise(absences = length(which(absent)),
            eligibles = length(absent)) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100)
y <- ab_panel %>%
  # filter(oracle_number %in% matched$oracle_number) %>%
  filter(!unidade %in% sort(irs$unidade)) %>%
  summarise(days_since = 'Never',
            absences = length(which(absent)),
            eligibles = n()) %>%
  mutate(absenteeism_rate = absences / eligibles * 100)
x <- bind_rows(x, y)
x$days_since <- factor(x$days_since,
                       levels = unique(c('Never',
                                  'Before',
                                  sort(unique(x$days_since)))))
cols <- c('red', rep('blue', nrow(x) - 1))
ggplot(data = x,
       aes(x = days_since,
           y = absenteeism_rate)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           fill = cols) +
  theme_maragra() +
  labs(x = 'Days since IRS',
       y = 'Absenteeism rate',
       title = 'Raw absenteeism as a function of days since IRS') +
  geom_label(aes(label = paste0(round(absenteeism_rate, digits = 2), '%')))
```

## Regression discontinuity analysis {.allowframebreaks}

\vspace{2mm}

All absenteeism:

\begin{tiny}

```{r}
kable(ors) 
```

\end{tiny}

\vfill
\newpage

```{r}
visualize_ors(ors_object = ors)
```

\vfill
\newpage

Sick absenteeism only:

\begin{tiny}

```{r}
kable(ors_sick) 
```

\end{tiny}

# Discussion

## General

- 30-50% reduction in absenteeism in the 6 months after IRS, when administered during malaria season.

## Implications

- Reduction is large and significant.
- Depending on detailed cost data (pending), IRS may be effective even from a purely financial point of view.

## Limitations

- No analysis yet of different worker types (agricultural vs. industrial).
- Have not yet ventured at all into side-analyses (effect on employment, tonnage, etc.). 
- Sick absenteeism seems to track absenteeism poorly: lack of clarity regarding pathways.

## Thank you

Your "pre-publication peer review" comments are appreciated:

**Email:** joe@economicsofmalaria.com

**Presentation:** [economicsofmalaria.com/ihmt](http://economicsofmalaria.com/ihmt)

## References {.allowframebreaks}