### 2018-01-03 inst/rmd/04_discussion.Rmd

  
# Discussion

Overall stuff

## Limitations

## Implications