### 2018-04-18 inst/rmd/04_discussion.Rmd

  
# Discussion

\todo{This entire section will not be done until methods/results are finalized.}

- Overview of findings

- How this contributes to the literature

- Implications for policy and for businesses

## Limitations

- No analysis yet of different worker types (agricultural vs. industrial).
- Have not yet ventured at all into side-analyses (effect on employment, tonnage, etc.).
- Sick absenteeism seems to track absenteeism poorly: lack of clarity regarding pathways.
- Large sample size, but all from same place: questionable generalizability.