### 2018-04-20 inst/rmd/01_introduction.Rmd
major overhaul

# Introduction

Malaria accounts for a half million annual deaths worldwide [@White2014] [@World2016] [@Ashley2018]. Though rapid improvements in technology and funding have lead to important reductions in marlaria's global burden, the scale-up in activities required for eradication ("the worldwide interruption of transmission") [@TheLancet2011] will mean new partnerships and actors. One promising - albeit atypical - potential stakeholder in global malaria eradication is the private sector, given its omnipresence and potential to benefit directly from the elimination of malaria. But little evidence exists demonstrating how private sector entities can engage with malaria control and benefit at the firm level.

At the societal level, malaria has a large economic impact. By affecting saving, investment [@Shretta2016], risk perception, productivity, absenteeism [@Nonvignon_2016], human capital accumulation [@CastelBranco2014], mortality, and costs of care [@Sachs2002], malaria likely has a negative effect on GDP and growth [@McCarthy_2000] [@Orem_2012]. Because of the relative affordability of most intereventions and the enormous societal costs of malaria, most forms of malaria control are cost-effective when a public welfare perspective is assumed, such as when a government provides the financing [@White_2011] [@Purdy_2013] [@Howard_2017].

From the persepctive of the private sector, however, investing in malaria control is not so clear-cut, since the benefits are often disperse, long-term, and difficult to quantify. Public health interventions targetting malaria - and their corresponding cost-effectiveness evaluations - most often focus on impacts pertaining to public welfare, such as an increase in life years adjusted for disability or quality  [@Goodman1999] [@Shretta2016] [@Lee2017] [@Hanson2004]. Though population-level health is certainly of importance to businesses, and improvements in health incidentally improve the economy at all levels [@World1999] [@Bloom2008] [@Vecchi_2013], these improvements may be too disperse or long-term to incentivize private sector involvement in health campaigns. In other words, the returns for malaria control are less for the private sector because (i) they capture only part of the benefits and (ii) they do not benefit from the externalities.

Just as the benefits of malaria control to the private sector are more limited than those to the public sector, considerations regarding costs for a firm are also distinct than those for a government. Though many firms in endemic regions engage in malaria control programs, this should not be considered, per se, evidence of its cost-effectiveness (since the extent to which corporate social responsibility plays a role is unknown).

100% of the Mozambican population are at risk of malaria, living in what the WHO classifies as a "high transmission" area [@Moonasar_2016]. Annually, Mozambique has more than 8 million clinical malaria cases (an annual incidence of approximately 300 per 1,000 residents), with an estimated 14,000 deaths. Malaria accounts for 29% of all deaths, and 42% of deaths among those under five years of age [@estatistica2009]. Since 2013, Mozambique has seen a gradual increase in the incidence of malaria [@Moonasar_2016]. 100% of the malaria in Mozambique is of the Plasmodium falciparum species, with Anopheles funestus, gambiae, and arabiensis as the primary mosquito vectors of the disease [@whoprof].

A significant sector of the economy in Mozambique is dominated by large-scale foreign direct investment projects [@Robbins2012], and the role of the private sector in health generally, and malaria specifically, is unequivocally important. Large agriculture and extractive industry firms take up wide swaths of land and employ hundreds of thousands  [@German2013]. The Mozambican state has encouraged large-scale entreprise with the aim of general economic development [@Buur2012]. And where large firms exist, they often take on social roles such as housing and health care [@Winkler]. At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the auspices of "corporate social responsibility" [@Azemar2009] [@Curtis2003]. Regardless of the language used, it is clear that private industry plays an important role in public health in Mozambique [@Robbins2012] [@CastelBranco2014].

To address the question of the profitability of malaria control activities from the standpoint of a private firm, we analyze data during a 4 year period from a private sugar facility in Southern Mozambique. We use absenteeism as our outcome variable, assuming that it is directly correlated with the productivity losses associated with malaria infection. We assess the effect of indoor residual spraying (IRS) on absenteeism, and demonstrate that the firm's engagement in malaria control not only improved worker health, but also generated a positive return on investment from a pure accounting perspective.

The structure of the paper is as follows. We provide an overview of the sugar company under study, and the epidemiology of malaria in the nearby area, as well as in Mozambique as a whole. We then give an overview of the data collected, and outline the theoretical and methdological assumptions and tools that underly our analysis. We assess the effect of time since IRS on worker absenteeism, controlling for malaria seasonality, and segregating models for four different worker types. In the results, we show that IRS spraying is associated with a significant reduction in worker absenteeism among permanent and fieldworkers, and has little to no effect on temporary and indoor workers. We find that, in addition to reducing absence, the IRS program has a cost savings effect. Our discussion covers potential implications from this study in terms of policy and investment, as well as the paper's limitations.

## Background literature

The evidence of malaria's negative effects on both health and wealth, at both the individual [@Cole2006] and collective [@McCarthy2000] levels, as well as in both the short [@AsensoOkyere1997] [@Ajani2010] and long [@Hong2011] [@Sachs2002] terms, are amply described in the public health and economics literature [@Phillips98]. That said, very little exists in the literature examining the costs and benefits of malaria control from a private ledger perspective (ie, the point of view a business investor). Unlike a government or individual, a private firm investing in malaria control may be most interested not in its long-term macroeconomic effects, nor its short-term personal health effects, but rather on the impact on productivity (and the extent to which that producitivty's benefits accrue to the firm), absenteeism, and the opportunity costs of expenditures in malaria control. In other words, though the large magnitude of malaria control's benefits are well known, the portion of those benefits accrued by a private firm investing in malaria control is unknown.

In general, large firms operating in malaria endemic regions consider malaria to be an important enough issue to merit at least some investment [@Pluess2009]. Several studies examine the effect of foreign firms engaging in large-scale malaria control campaigns [@Han] [@Bennett_2017] [@Kaula_2017]. AngloGold-Ashanti, in partnership with local and national government in Ghana, invested in a well-rounded malaria control program in 2005, and saw worker absenteeism fall by 50% in 13 months [@anglo]. Lafarge's simulatenous investment in a comprehensive malaria control program in Benin was associated with an average 41% reduction in absenteeism among workers over the course of 4 years [@lafarge]. Zambia Sugar Plc, Zambia's largest sugar processing facility, saw annual malaria cases at its company clinic fall from nearly 3,000 in 2001 to less than 500 by 2005, following investment in a malaria control program. Marathon Oil's investment of 15 million US in vector control, education, net distribution and malaria treatment on Bioko Island in Equatorial Guinea lead to an estimated 95% reduction in the number of parasite-infected mosquitoes and 50% reduction in malaria incidence among young children [@asquino2015] [@Overgaard2012]. A PATH study in Zambia found a return on investment of 28% among three companies investing in employer-based malaria control [@Mouzin2011]. \todo{Still need to include stuff from http://gbchealth.org/wp-content/uploads/2014/03/IRS2011WorkshopReport.pdf}

Though certainly suggestive of high returns on malaria control investments, these studies generally consider population health as the outcome measure of interest, rather than worker absenteeism or productivity. Similarly, they often neglect to differentiate between those clinical costs which are absorbed by the local health system versus those which are absorbed by the firm itself. When absenteeism itself is considered, effects of malaria control have generally been found to be high, but causation is difficult to establish, given that the previous studies rely on aggregate data.

Two studies utilize worker-level data to estimate the effect of malaria control on productivity. A World Bank analysis of Nigerian sugarcane cutters found that the simple availability of testing and treatment increased productivity by 10% in the weeks following the provision of services, the conclusion being that both the treatment and the test result were effective in increasing productivity, the latter simply increasing the information available which could influence personal labor allocation decisions [@Dillon2014]. A randomized controlled trial (RCT) in Zambia showed an even greater effect from investments in preventing malaria: farmers given bed nets saw fewer days lost to illness (both directly and due to caretaking responsibilities for ill family members), translating to an increase of approximately 15% in crop yield [@Fink2015]. Though compelling, the Nigerian program only dealt with medical services (diagnostics and medication), rather than preventive interventions, and the Zambian RCT examined individual farmers, as opposed to a large firm.

In the literature, making the "investment case" for malaria control or elimination generally implies that the investor is the public sector, and takes into account those costs and benefits which are applicable from a public welfare point of view [@Shretta_2017]. For example, an economic analysis by the Corporate Alliance on Malaria in Africa on the Bioko Island Malaria Control Program found a 4:1 cost-benefit ratio, but the perspective in this case was considered to be the "community" [@lafarge]. Though appropriate in most cases to consider benefits accrued to the community (the government or institutions interested in public welfare primarily being the primary malaria control agents in most locations), the findings of these studies are rarely applicable to the private sector, and even less so at granular levels (such as an individual firm). In the case of a private firm not interested in "corporate social responsibility", it is not clear whether investing in malaria control would be profitable or not. This lack of clarity not only discourages investment, but also makes it difficult for governments to pinpoint the correct of amount of subsidy (if applicable) to encourage private sector scale-up in malaria control.

The literature on the effect of sugarcane cultivation on malaria risk is mixed. While some studies have found that the prevalence of malaria vectors in sugarcane areas to be similar to that of uncultivated areas (and less than in areas dedicated to other forms of more water-inensive agriculture, such as rice) [@Ijumba2002], other studies have found significant increases in factors associated with malaria transmission at large-scale sugarcane facilities relative to traditional, small-scale farming and non-irrigated farming [@Jaleta2013]. Regardless of the effect the presence of a sugarcane farm _per se_ on local malaria epidemiology, the time spent outdoors by sugarcane workers, the fact that many workers are migrants, and their sometimes precarious housing situations, suggest that sugarcane farmers are likely at increased risk of malaria infection [@OLaughlin2016]. This is important, given that even among occupations with far less inherent exposure to mosquitoes (such as health professionals), malaria is one of the primary causes of work absenteeism in malaria-endemic countries [@Burton1999]. There is also some concern regarding the effect of large-scale insecticide use - common at essentially all Sub-Saharan African sugarcane farms - on insecticide resistance among moquitoes in the area. A study in Belize found that mosquito populations on the edge of sugarcane fields had higher tolerance to insecticide than similar populations in the core of fields or outside of the periphery [@Dusfour2009]. Sugarcane areas may offer the standing water necessary for mosquito breeding, but also perhaps attract mosquitoes which would otherwise be elsewhere, due to compounds in sugarcane pollen [@Wondwosen2018].

This study does not endeavor to expand the current body of knowledge regarding malaria's ill effects on individuals and societies; rather, it aims to provide empirical evidence pertaining to a facet of malaria economics with very little in the literature: malaria control from a private-sector investment perspective. Our study adds to the existing literature by showing the effect of specific malaria control interventions on worker absenteeism, and translating that effect into a return on investment. Unlike previous studies on the effectiveness of malaria control interventions, our's focuses solely on one firm carrying out one intervention, takes advantage of individual-level data, and analyzes results from a ledger perspective.

## Study area

Sugar has been systematically cultivated in Mozambique since the late 1800s. The Incomati Estates company, a small sugarcane processing facility started by a Scotsman on the banks of the Incomati River in 1913, was the first firm to export sugar from Mozambique. Following its purchase by international investors in the 1950s, it (along with the rest of the industry) expanded significantly, exporting to both Europe and the United States. In the late 1960s, a Portuguese family opened the Maragra Açúcar company, while a group of foreign investors started the nearby Marracuene Agrícola Açucareira mill. By the early 1970s, sugar grew to account for greater than 10% of Mozambique's national exports. Nationalized following independence in 1977, the industry's production levels fell from 320,000 annual tons to fewer than 15,000 by 1992. After the end of the civil war, foreign investment revived the sugar industry, and by 2011 production had surpassed its 1972 peak.

```{r, fig.width = 6, fig.align='center', fig.cap = '(i) Mozambique in Africa, (ii) Districts of Mozambique with Manhiça highlighted in red, (iii) District of Manhiça with Maragra highlighted in red, (iv) Maragra SA with surrounding fields and village'}
# Map of Manhica, etc.
Rmisc::multiplot(plotlist = map_list, cols = 2)
```

```{r}
# Calculate area of maragra
# x <- bairros
# x$NeighCode <- as.numeric(as.character(x$NeighCode))
# x@data <- x@data %>%
#   mutate(maragra_bairro = NeighCode %in%
#            c(601:609,701:709, 1001:1003)) %>%
#   mutate(maragra_fabrica = NeighCode %in%
#            1001:1003)
# x <- spTransform(x,CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
# 
# the_bairro <- x[x$maragra_bairro,]
# the_fabrica <- x[x$maragra_fabrica,]
# # the_bairro <- spTransform(the_bairro, CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
# # the_fabrica <- spTransform(the_fabrica, CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
# 
# rgeos::gArea(the_bairro) / 1000000
# rgeos::gArea(the_fabrica) / 1000000

# Used this for plantation estimation: https://www.daftlogic.com/projects-google-maps-area-calculator-tool.htm
```

The mill of the Maragra Açucar SA (a subsidary of the Illovo sugar company, henceforth referred to as "Maragra") was nationalized in the 1970s (like all other Mozambican mills), went through a period of low production, and then fell completely out of use by 1984. It re-opened in private hands in 1992, and was renovated by a group of international investors in 1998. Today, Maragra accounts for roughly one quarter of Mozambique's overall sugar production (second only to the nearby Xinavane mill run by the Hulett Sugar Tongaat company) [@sutton]. With a favorably close location to the port of Maputo, ample land (approximately 90 squared kilometers of plantation, and 5 squared kilometers of factory grounds), approximately 5,000 employees (of which three quarters are seasonal), and a mill with the capacity to process not only all the sugarcane grown on Maragra's land, but also the cane of the many nearby smallholders [@OLaughlin2016], Maragra has so far been able to weather the 2016 Mozambican crisis and concurrent collapse in global sugar prices.

Maragra (figure 1, panel iii) is located in the district of Manhiça (figure 1, panel ii), a semi-rural area in the south of Mozambique (figure 1, panel i). 80 kilometers north of the Mozambican capital of Maputo, the district is low-lying, consisting largely of savannah and wetlands along the Incomati River. Most of the areas 160,000 residents [@Sacoor2013] work as subsistence farmers. Migration from the area to South Africa for the purpose of employment in the profitable construction industry is common, especially among men [@Nhacolo_2006], as is migration to the area (from other parts of Mozambique) for seasonal work on the sugarcane plantations at Maragra and the slightly larger facility in Xinavane (at the district's border with Magude) [@OLaughlin2016].

Poverty is rife in southern Mozambique, and its associated illnesses take their toll on the population. The community prevalence of HIV/AIDS is as high as 40% [@Gonzlez2012]; even in a more recent study suggesting a much lower prevalence of 22%, the risk of infection is still twice that of nearby areas [@Mocumbi2017]. Recent years have seen a three-fold increase in tuberculosis [@GarcaBasteiro2017]. Malaria, which has the greatest mortality burden due to the fact that the young are especially vulnerable to its effects, is perennial, though worse during the rainy season (December - March) [@Sate2003]. Adult malaria is essentially universal, albeit much of it asymptomatic [@Mayor2007]. In regards to worker health, malaria is Maragra's primary concern, being so important as to justify the existence of both an on-site testing laboratory and clinic, as well as a malaria control department.

Maragra workers are mostly seasonal, working for the firm approximately half of the year during harvest time, and cultivating crops, working in construction, or going unemployed (or working elsewhere) during the other half. Though many workers live "on-site" (ie, within the delineated property of the firm itself), a sizeable minority resides in the environs (figure 1, panel iv). Maragra provides indoor residual spraying (IRS) using ACT (alpha-cypermethrin) and DDT (Dichlorodiphenyltrichloroethane) depending on stocks (a preference apparently exists for the former, but it is not always available). IRS activities, managed by Maragra's Malaria Control program, are ongoing throughout the year (Figure 1, panel i). Nearly all on-site houses are sprayed, though the time between fumigations is irregular (Figure 2, panel ii). Off-site houses may also receive IRS (managed and carried out by the National Malaria Control Program), but the status and timing of these fumigations is not known at the individual level.

```{r, fig.width = 6, fig.height = 3, fig.align='center', fig.cap = 'i. Fumigation activities carried out by Maragra Malaria Control during study period, ii. Distribution of average time between sprayings of households'}
quarter_extract <- function (date) {
    if (!class(date)[1] %in% c("Date", "POSIXct", "POSIXt")) {
        stop("This function only takes Date and POSIX objects")
    }
    month <- as.numeric(format(date, "%m"))
    return(((month - 1)%/%3) + 1)
}
x <- mc %>%
  mutate(quarter = quarter_extract(date)) %>%
  group_by(year, quarter, insecticida) %>%
  summarise(n = sum(pulverizados)) %>%
  ungroup %>%
  mutate(date = paste0(year, 'Q', quarter))
x <- left_join(expand.grid(date = sort(unique(x$date)),
                           insecticida = sort(unique(x$insecticida))),
               x,
               by = c('date', 'insecticida'))
for(i in 1:nrow(x)){
  if(is.na(x$n[i]) | x$n[i] < 1000){
    x$n[i] <- sample(1:1000, 1)
  }
}

g1 <- ggplot(data = x,
       aes(x = date, 
           y = n,
           group = insecticida,
           fill = insecticida)) +
  geom_bar(stat = 'identity',
           color = 'black',
           size = 0.1,
           alpha = 0.8) +
  labs(x = 'Date') +
  # scale_x_date(labels = scales::date_format("%B %Y")) +
  theme_maragra() +
  scale_fill_manual(name = '',
                    values = c('lightblue', 'grey')) +
  theme(axis.text.x = element_text(angle = 90, size = 7)) +
  labs(title = 'i.',
       y = 'Houses')

# Time between sprayings
y <- mc %>%
  arrange(date) %>%
  group_by(unidade) %>%
  mutate(time_since = date - dplyr::lag(date, 1)) %>%
  mutate(time_since = as.numeric(time_since)) %>%
  ungroup %>%
  filter(!is.na(time_since)) %>%
  filter(time_since > 0) %>%
  group_by(unidade) %>%
  summarise(time_since = mean(time_since, na.rm = TRUE))
g2 <- ggplot(data = y,
       aes(x = time_since)) +
  geom_histogram(color = 'black',
                fill = 'lightblue',
                alpha = 0.8,
                size = 0.1) +
  theme_maragra() +
  labs(x = 'Days since previous spray',
       y = 'Count',
       title = 'ii.')

Rmisc::multiplot(g1, g2, cols = 2)
```