### 2018-04-20 inst/rmd/04_discussion.Rmd

  
# Discussion

\todo{This entire section will not be done until methods/results are finalized.}

- Overview of findings
- How this contributes to the literature
- Implications for policy and for businesses
- Spillover effects !
- How absenteeism might be _underestimating_ true effect, since at the margin a sick worker might go to work (ie, not be absent), but be less productive

## Limitations

- No analysis yet of different worker types (agricultural vs. industrial).
- Have not yet ventured at all into side-analyses (effect on employment, tonnage, etc.).
- Sick absenteeism seems to track absenteeism poorly: lack of clarity regarding pathways.
- Large sample size, but all from same place: questionable generalizability.