### 2018-05-15 inst/rmd/02_methods.Rmd
major changes

# Methods

## Budget info

```{r}
costs_itemized <- maragra::costs_itemized
costs_raw <- maragra::costs_raw
costs <- maragra::costs
# Annual operating budget
aob <- round(mean(costs$usd), digits = 2)
print(paste0('Average annual operating budget of ', aob))

plot_data <- costs_itemized %>%
  filter(!is.na(Account)) %>%
  mutate(area = Account) %>%
  mutate(area = ifelse(grepl('WAGES', area), 'Human resources', area)) %>%
  mutate(area = ifelse(area %in% c('LUBRICANTS', 'TYRES & TUBES',
                                   'VEHICLE MAINTENANCE', 'DIESEL'),
                       'Vehicles', area)) %>%
  mutate(area = ifelse(area %in% c('PENSION FUND CONTRIBUTIONS',
                                   'SALARIES BASIC', 'HOUSING ALLOWANCE',
                                   'LEAVE PAY PROVISION'), 'Human resources', area)) %>%
  mutate(area = ifelse(area %in% toupper(c('Equipment maintenance',
                                   'Protective clothing',
                                   'Tools',
                                   'Consumables')),
                       'Equipment', area)) %>%
  mutate(area = ifelse(area %in% c('ENVIRONMENTAL MONITORING',
                                   'TRAVEL (FOREIGN)'),
                       'Other', area)) %>%
  group_by(period, area) %>%
  summarise(total = sum(usd, na.rm = TRUE)) %>%
  ungroup %>%
  mutate(area = tolower(area)) %>%
  mutate(area = Hmisc::capitalize(area))

ggplot(data = plot_data,
             aes(x = period,
                 y = total,
                 group = area,
                 fill = area)) +
  geom_bar(stat = 'identity',
           position = 'dodge') +
  scale_fill_manual(name = '',
                    values = colorRampPalette(brewer.pal(n = 9,
                                                         name = 'Set1'))(length(unique(plot_data$area)))) +
  theme_maragra() +
  labs(x = '',
       y = 'USD')

fumigations <- mc %>%
  filter(date >= '2013-01-01',
         date <= '2016-12-31')
n_fumigations <- nrow(fumigations)
n_buildings <- sum(fumigations$casas_cobertas, na.rm = TRUE)
unique_agregados <- length(unique(fumigations$unidade))
people_sprayed <- length(unique(model_data$oracle_number[model_data$days_since != 'Never']))
people_observed <- length(unique(model_data$oracle_number))
pr <- function(x){prettyNum(x, big.mark = ',')}

print('Assuming linearity in costs, the program spends approximately xxx per building sprayed.') 
aob / (n_buildings / 4)

# print('With each agregado containing an average of 2.2 workers.')

x = ab_panel %>%
  mutate(year = as.numeric(format(date, '%Y')),
         month = as.numeric(format(date, '%m'))) %>%
  group_by(year, month) %>%
  summarise(absences = length(which(absent))) %>%
  ungroup %>%
  left_join(clinic_agg %>% group_by(year, month) %>% 
              summarise(positive = sum(positive))) %>%
  mutate(percent_clinic = positive / absences * 100) %>%
  mutate(n_ab_for_clinic = absences / positive)

mean(x$n_ab_for_clinic, na.rm = TRUE)
```

```{r, fig.cap = 'Kernel density surface of average all-time herd protection score for Maragra worker neighborhoods (relative to highest protection)', fig.align = 'center'}
# Fortify bairros
library(maptools)
library(rgeos)
border <- gConvexHull(bairros_maragra_bairro)
border <- SpatialPolygonsDataFrame(Sr = border,
                                   data = data.frame(a = 1))
# border <- unionSpatialPolygons(bairros_maragra_bairro, 
#                           IDs = rep(1, nrow(bairros_maragra_bairro)),
#                           threshold = 9^9)
border_fortified <- 
  broom::tidy(border,
              regions = 'Id')

# Create average herd risk score
x <- model_data %>%
  group_by(lng = longitude_aura,
           lat = latitude_aura) %>%
  summarise(herd = mean(herd, na.rm = TRUE)) %>%
  ungroup %>%
  filter(!is.na(lng))
x_sp <- x
coordinates(x_sp) <- ~lng+lat   
proj4string(x_sp) <- proj4string(border)
overs <- sp::over(x_sp, border)
x <- x[!is.na(overs),]
# ggplot() +
#   geom_point(data = x,
#        aes(x = lng,
#            y = lat,
#            color = herd)) +
#   geom_polygon(data = border_fortified,
#                aes(x = long,
#                    y = lat,
#                    group = group),
#                fill = NA,
#                color = 'black') +
#   ggthemes::theme_map() +
#   scale_color_continuous(name = 'Herd\nprotection\nscore',
#                          low = 'orange',
#                          high = 'blue')

df_grid <- expand.grid(lng = seq(bbox(border)[1,1],
                                 bbox(border)[1,2],
                                 by = 0.0008),
                       lat = seq(bbox(border)[2,1],
                                 bbox(border)[2,2],
                                 by = 0.0008),
                       herd = NA)
df_grid$latitude <- df_grid$lat
df_grid$longitude <- df_grid$lng
coordinates(df_grid) <- ~longitude+latitude

# Go through each row of df_grid, getting the 
# weighted mean irs score for that point
# and putting a color into df_grid
if('df_grid.RData' %in% dir()){
  load('df_grid.RData')
} else {
  for (i in 1:nrow(df_grid)){
  # Get distance from every point in x_sp
  distances <- spDistsN1(pts = x_sp,
                        pt = df_grid[i,],
                        longlat = TRUE)
  # Define which are acceptably close
  close_enough <- which(distances <= 50)
  # Get an herd score
  herd <- stats::weighted.mean(x = x_sp$herd[close_enough],
                       w = (1 / distances[close_enough]) ^2,
                       na.rm = TRUE)
  # Assign irs to the dataframe
  df_grid$herd[i] <- herd

}
save('df_grid', file = 'df_grid.RData')
}

library(raster)
# Convert df_grid to raster
temp <- df_grid@data %>% arrange(lng, lat)
r <- rasterFromXYZ(temp[, c('lng', 'lat', 'herd')])

proj4string(df_grid) <- proj4string(border)
x <- over(df_grid, polygons(border))
df_grid_small <- df_grid[!is.na(x),]
temp <- df_grid_small@data %>% arrange(lng, lat)
r <- rasterFromXYZ(temp[, c('lng', 'lat', 'herd')])
values(r) <- values(r) / max(values(r), na.rm = TRUE) * 100
plot(border, col = NA, border = NA)
plot(r, add = TRUE)
# plot(border, add = TRUE)
plot(bairros, add = TRUE)
```

```{r}
x <- model_data %>%
  group_by(group) %>%
  summarise(avg = mean(herd, na.rm = TRUE),
            avg_ideal = mean(herd_ideal, na.rm = TRUE)) %>%
  ungroup %>%
  mutate(avg = avg / max(avg) * 100,
            avg_ideal = avg_ideal / max(avg_ideal) * 100)
```

### Econometric model

Our model specification is as follows

$$
\hat{Y_{it}} = \hat{\beta}_{0} +  \hat{\beta}_{1}\text{Season}_{t} * (\hat{\beta}_2{IRS_{it}}*\hat{\beta}_3{IRStime_{it}}) +  \hat{\beta}_4{Herd_{it}} +  \alpha_i + \delta_t + \upsilon_{it}
$$

$\hat{Y}_{it}$ is the rate of absence. $\beta_{1}$ is the binary "season" variable, imputed from overall district clinical incidence. Our intervention is whether the residence of the worker in question was treated in the last year, and, if so, the time since treatment, represented, respectively, by $\beta_{2}$ and $\beta_{3}$. $\beta_{4}$ **represents the distance-weighted herd protection score**. $\alpha_i$ represents the time invariant worker fixed effects, and $\delta_i$ represents the fixed effect of the particular malaria season. $\upsilon$ is the error term.

Our formula for return on investment can be described in a straightforward fashion...

\begin{center}
$R = \dfrac{P_{w} - S_{wa} - S_{wc}}{P_{w}}$

\end{center}

...where $R$ is the return on investment, $P$ is the malaria control program's total operating cost, $w$ refers to costs at the per-worker level, $a$ refers to savings through avoided absences, and $ c $ refers to savings through avoided clinical encounters. We define the malaria control program as "profitable" from an investment standpoint if ROI is greater than 100%, ie if the savings associated with the estimated effect of IRS is greater than the costs of the program's administration.

\newpage