### 2019-02-17 inst/rmd/final/final.Rmd
minor update

---
title: "Final analysis for paper"
subtitle: "Brew, Pradhan, Sicuri"
date: ''
output: 
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# csl: journal-of-health-economics.csl

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
library(raster)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 4,
               fig.width = 5,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               dev = "cairo_pdf",
               fig.pos="!h",
               fig.align = 'center')
opts_knit$set(width=75)
options(xtable.comment = FALSE)
```

```{r, eval = TRUE}
source('prepare_data.R')
```

# Manuscript draft

Draft of the paper (\href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}

# Descriptive overview

## Herd-only protection score

```{r}
pq <- quantile(final_data$herd_protection)
x <- final_data %>%
  # mutate(protection_cat = protection_var) %>%
  mutate(protection_cat = ifelse(herd_protection <= pq[2],
                                 'Protection 1',
                                 ifelse(herd_protection <= pq[3],
                                        'Protection 2',
                                        ifelse(herd_protection <= pq[4],
                                               'Protection 3',
                                               ifelse(herd_protection <= pq[5],
                                                      'Protection 4',
                                                      NA))))) %>%
  group_by(protection_cat) %>%
  summarise(ab_rate = length(which(absent))/ n() * 100)

names(x) <- c('Quantile', 'Absenteeism rate')
kable(x)
```

```{r}
ggplot(data = x,
       aes(x = `Quantile`,
           y = `Absenteeism rate`)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(x = 'Protection score quantile')
```

\newpage

## Individual-only protection score

```{r}
pq <- quantile(final_data$protection,
               seq(0, 1, 0.1))
pq <- sort(unique(pq))
x <- final_data %>%
  # mutate(protection_cat = protection_var) %>%
  mutate(protection_cat = ifelse(protection <= pq[2],
                                 'Protection 1',
                                 ifelse(protection <= pq[3],
                                        'Protection 2',
                                        ifelse(protection <= pq[4],
                                               'Protection 3',
                                               ifelse(protection <= (pq[5] + 0.1),
                                                      'Protection 4',
                                                      NA))))) %>%
  group_by(protection_cat) %>%
  summarise(ab_rate = length(which(absent))/ n() * 100)

names(x) <- c('Quantile', 'Absenteeism rate')
kable(x)
```

```{r}
ggplot(data = x,
       aes(x = `Quantile`,
           y = `Absenteeism rate`)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(x = 'Protection score quantile')
```

\newpage

## Overall protection score

```{r}
pq <- quantile(final_data$protection_var)
x <- final_data %>%
  # mutate(protection_cat = protection_var) %>%
  mutate(protection_cat = ifelse(protection_var <= pq[2],
                                 'Protection 1',
                                 ifelse(protection_var <= pq[3],
                                        'Protection 2',
                                        ifelse(protection_var <= pq[4],
                                               'Protection 3',
                                               ifelse(protection_var <= pq[5],
                                                      'Protection 4',
                                                      NA))))) %>%
  group_by(protection_cat, group) %>%
  summarise(ab_rate = length(which(absent))/ n() * 100) %>%
  ungroup

names(x) <- c('Quantile', 'Group', 'Absenteeism rate')
kable(x)
```

```{r}
ggplot(data = x,
       aes(x = `Quantile`,
           y = `Absenteeism rate`)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(x = 'Protection score quantile') +
  facet_wrap(~Group)
```

\newpage

## Precipitation lag

```{r}
rq <- quantile(final_data$rain_var, na.rm = TRUE)
x <- final_data %>%
  # mutate(protection_cat = protection_var) %>%
  mutate(protection_cat = ifelse(rain_var <= rq[2],
                                 'Protection 1',
                                 ifelse(rain_var <= rq[3],
                                        'Protection 2',
                                        ifelse(rain_var <= rq[4],
                                               'Protection 3',
                                               ifelse(rain_var <= rq[5],
                                                      'Protection 4',
                                                      NA))))) %>%
  group_by(protection_cat, group) %>%
  summarise(ab_rate = length(which(absent))/ n() * 100) %>%
  ungroup %>%
  filter(!is.na(protection_cat))

names(x) <- c('Quantile', 'Group', 'Absenteeism rate')
kable(x)
```

```{r}
ggplot(data = x,
       aes(x = `Quantile`,
           y = `Absenteeism rate`)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(x = 'Lagged precipitation score quantile') +
  facet_wrap(~Group)
```

\newpage

# Results

## Regression table
The below table shows the results of the model devised thus far.

```{r, results = 'asis'}
# make_final_table(model = final_model, the_caption = "Model results", type = 'latex', multiplier = 10000)
```

```{r}
summary(final_model)
```

## Intepretation

The below charts show predicted absenteeism rates at different rain and protection levels.

```{r}
fake <- expand.grid(#group = sort(unique(monthly$group)),
  rain_var = rq,
  protection_var = pq)#,
fake$predicted <- fake$predicted_upr <- fake$predicted_lwr<- NA

fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, fake, interval = 'confidence'))-1
# predictions <- data.frame(predictions)
# # Add the fixed effects
# fixed_effects <- fe$effect

fake$predicted<- predictions[,1]
fake$predicted_upr <- predictions[,3]
fake$predicted_lwr <- predictions[,2]
```

### Precipitation's effect

```{r}
ggplot(data = fake %>% filter(protection_var == pq[1]),
       aes(x = rain_var,
           y = predicted)) +
  geom_point() +
  geom_line() +
  geom_ribbon(aes(ymin = predicted_lwr,
                  ymax = predicted_upr),
              alpha = 0.4,
              fill = 'darkorange') +
  labs(y = 'Predicted absenteeism rate',
       x = '1-3 month precipitation lag',
       title = 'Lagged precipitation and absenteeism')
```

### IRS protection's effect

```{r}
ggplot(data = fake %>% filter(rain_var == rq[1]),
       aes(x = protection_var,
           y = predicted)) +
  geom_point() +
  geom_line() +
  geom_ribbon(aes(ymin = predicted_lwr,
                  ymax = predicted_upr),
              alpha = 0.4,
              fill = 'darkorange') +
  labs(y = 'Predicted absenteeism rate',
       x = 'Protection index',
       title = 'IRS-afforded protection and absenteeism')
```

### Simulations

```{r}
sim_data <- simulations
sim_data$idx <- sim_data$oracle_number
fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, sim_data, interval = 'confidence'))-1
predictions <- data.frame(predictions)
# Add the fixed effects
fixed_effects <- left_join(sim_data,
                           data.frame(fe) %>% dplyr::select(idx, effect) %>%
                             mutate(idx = as.character(idx))) %>%
  .$effect
fixed_effects <- as.numeric(fixed_effects)

predictions$fit <- predictions$fit + fixed_effects
predictions$lwr <- predictions$lwr + fixed_effects
predictions$upr <- predictions$upr + fixed_effects

# combine other info
predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy)) %>%
  mutate(absent = as.numeric(absent)) %>%
  filter(!is.na(fit))

simple <- predictions %>%
  group_by(strategy) %>%
  summarise(n = n(),
            actual_absences = sum(absent),
            predicted_absences = sum(fit, na.rm = TRUE),
            actual = mean(absent, na.rm = TRUE),
            predicted = mean(fit, na.rm = TRUE)) %>%
  dplyr::rename(value = predicted) %>%
  dplyr::select(strategy, value, predicted_absences)

kable(simple)
```

```{r}
ggplot(data = simple,
       aes(x = strategy,
           y = value)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Absenteeism rate',
       x = 'Strategy') +
  geom_label(aes(label = round(value, digits = 3)),
             nudge_y = -0.005)
```

```{r}
ggplot(data = simple,
       aes(x = strategy,
           y = predicted_absences)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Absenteeism rate',
       x = 'Strategy') 
```

# Costs

```{r}
x <- final_data %>%
  left_join(irs %>% dplyr::select(date, unidade, days_since))

out <- data.frame(
  strategy = c('Max',
               'Real',
               'Time-optimized',
               'Zero'),
  sprays = c(nrow(final_data) / 60,
             length(which(x$days_since == 0)),
             length(which(format(final_data$date, '%m-%d') == '12-11')),
             0
  )
) %>%
  mutate(sprays = sprays / length(unique(final_data$oracle_number)))

kable(out)
ggplot(data = out,
       aes(x = strategy,
           y = sprays)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Number of sprays per person',
       x = 'Strategy')

```

```{r}
# Introduction is really all that matters - specifying contribution, summarizing results
# Journal of Health Economics (1) and then Health Economics (2).
# Elisa: American Journal of Health Economics (from American Health Economic Society)
# Ensure that we explicitly explain these 21 days issue to rule out reverse causality.
# Large firms are big enough to capture externalities
# What is the effect of anti-malaria campaign on worker producitivity
# Data special
# Identification problems (no external spraying, no home spraying, unobserved govt - omitted variable bias dealt with through worker and time fixed effects.)
# Exploring herd protection through various functional forms
```

# Bibliography