### 2020-04-18 inst/rmd/final/final_september_2019.Rmd

---
title: "Final analysis for paper"
date: ''
subtitle: Brew, Pradhan, Sicuri
bibliography: bibliography.bib
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# csl: journal-of-health-economics.csl

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
library(raster)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               fig.height = 4,
               fig.width = 5,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               dev = "cairo_pdf",
               fig.pos="!h",
               fig.align = 'center')
opts_knit$set(width=75)
options(xtable.comment = FALSE)
```

```{r, eval = TRUE}
source('prepare_data.R')
```

# September 2019 stuff

## Table of observations

```{r}
library(table1)
message('We observed ', nrow(model_data), ' worker-days among ', length(unique(model_data$oracle_number)), ' unique workers. Overall absence rate was ', length(which(model_data$absent)))

# Table of those absences
vars <- list('field',
          'permanent_or_temporary',
           'sex')
table1::label(model_data$field) <- 'Worker location'
table1::label(model_data$permanent_or_temporary) <- 'Worker contract'
table1::label(model_data$sex) <- 'Worker sex'
model_data$year <- as.numeric(format(model_data$date, '%Y'))
table1::label(model_data$year) <- 'Year'

table1::table1(~field + permanent_or_temporary + sex | year , data = model_data)
```

## Table of absences

```{r}
calculate <- function(var){
  out <- 
  final_data %>%
    mutate(year = format(date, '%Y')) %>%
    # mutate(on_site = ifelse(on_site,
    #                                 # 'On site',
    #                                 # 'Off site'),
    #        rainy = ifelse(rainy,
    #                       'Rainy',
    #                       'Dry')) %>%
    # filter(!is.na(rainy)) %>%
  group_by_(var, 'year') %>%
  summarise(absences = length(which(absent)),
            eligibles = length(absent),
            workers = length(unique(oracle_number))) %>%
  ungroup %>%
  mutate(absenteeism_rate = absences / eligibles * 100) 
  names(out)[1] <- 'variable'
  out$val <- paste0(round(out$absenteeism_rate, digits = 1),
                    '%')#,
                    # ')',
                    # '\n',
                    # ' (workers:', out$workers,
                    # ')')
  out <- 
    out %>%
    dplyr::select(variable, year, val)

  out <- out %>%
    spread(key = year,
           value = val)
  out$variable <- as.character(out$variable)
  new_var <- #ifelse(var == 'on_site',
                    # 'Residence',
                    ifelse(var == 'rainy',
                           'Precipitation',
                           var)
  # )
  left <- data.frame(x = c(new_var, rep('', (nrow(out) - 1))))
  out <- cbind(left, out)
  names(out)[1:2] <- c('Variable', 'x')
  # out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
  return(out)
}
out_list <- list()
vars <- c(#'season', 
           'field',
          'permanent_or_temporary',
           'sex'#,
           # 'on_site',
           # 'rainy'
          )
for(i in 1:length(vars)){
  x <- calculate(var = vars[i])
  out_list[[i]] <- x
}
ab_table <- bind_rows(out_list)
ab_table$Variable <- Hmisc::capitalize(ab_table$Variable)
# ab_table$Variable[ab_table$Variable == 'Season'] <- 'Malaria season'
ab_table$Variable[ab_table$Variable == 'Field'] <- 'Worker type'
ab_table$Variable[ab_table$Variable == 'Permanent_or_temporary'] <- 'Contract'
ab_table$x <- Hmisc::capitalize(ab_table$x)
names(ab_table)[2] <- ''
ab_table$`2016`[is.na(ab_table$`2016`)] <- '0%'
kable(ab_table, format = 'html') %>% kable_styling(full_width = FALSE)
```

# Scenarios: policy simulations

## Cost of max strategy

```{r}
x = simulations %>%
  arrange(oracle_number, date) %>%
  mutate(first_day = protection == 1 & dplyr::lag(protection, 1) != 1) %>%
  filter(maragra_bairro) %>%
  group_by(strategy) %>%
  summarise(day1 = length(which(first_day)))
x

y <- maragra::costs_itemized %>% filter(area == 'Pesticides')
y <- y[3:6,]
sum(y$total)
```

## Breakdown

```{r}
sim_data <- simulations
sim_data$idx <- sim_data$oracle_number
fe <- getfe(final_model)
predictions <- exp(predict(final_model_lm, sim_data, interval = 'confidence'))-1
predictions <- data.frame(predictions)
# Add the fixed effects
fixed_effects <- left_join(sim_data,
                           data.frame(fe) %>% dplyr::select(idx, effect) %>%
                             mutate(idx = as.character(idx))) %>%
  .$effect
fixed_effects <- as.numeric(fixed_effects)

predictions$fit <- predictions$fit + fixed_effects
predictions$lwr <- predictions$lwr + fixed_effects
predictions$upr <- predictions$upr + fixed_effects

# combine other info
predictions <- predictions %>% bind_cols(sim_data %>% dplyr::select(date, absent, strategy, permanent_or_temporary)) %>%
  mutate(absent = as.numeric(absent)) %>%
  filter(!is.na(fit))

simple <- predictions %>%
  group_by(strategy) %>%
  summarise(n = n(),
            actual_absences = sum(absent),
            predicted_absences = sum(fit, na.rm = TRUE),
            actual = mean(absent, na.rm = TRUE),
            predicted = mean(fit, na.rm = TRUE)) %>%
  dplyr::rename(value = predicted) %>%
  dplyr::select(strategy, value, predicted_absences)

zero <- predictions %>%
  filter(strategy == 'Zero') %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  group_by(year) %>%
  summarise(n = n(),
            actual_absences = sum(absent),
            predicted_absences = sum(fit, na.rm = TRUE),
            actual = mean(absent, na.rm = TRUE),
            predicted = mean(fit, na.rm = TRUE)) %>%
  mutate(avoided = predicted_absences - actual_absences)
  # dplyr::rename(value = predicted) %>%
  # dplyr::select(permanent_or_temporary, value, predicted_absences)

```

```{r}
ggplot(data = simple,
       aes(x = strategy,
           y = value )) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Absenteeism rate',
       x = 'Strategy') +
  geom_label(aes(label = round(value, digits = 3)),
             nudge_y = -0.005) +
  theme_maragra() +
  theme(axis.text.x = element_text(size = 15))
```

```{r}
simple
ggplot(data = simple,
       aes(x = strategy,
           y = predicted_absences)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           color = 'black') +
  labs(y = 'Absenteeism rate',
       x = 'Strategy') 
```

# Manuscript draft

Draft of the paper \href{https://docs.google.com/document/d/1bUWRBCgVcgjSPHchIQxiTG8Vwv5hV1GLU4Tlu386sWA/edit#}{HERE}.

# Bibliography